---
title: 'Child Doc 1: Endpoint Phenotypic Data'
author: "Grace Rubin"
date: "`r Sys.Date()`"
output: html_document
editor_options:
  chunk_output_type: console
---

<!-- This is the child document containing data visualizations and statistical analyses for MFC2 Endpoint Phenotypic Data -->

```{r setup child doc 1, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  collapse = TRUE,
  fig.height = 3, 
  fig.width = 4, 
  fig.align = 'center'
)

# loading packages
library(ggpubr)
library(nlme)
library(tidyverse)
library(here)
library(knitr)
library(DT)
```

```{r Reading data, include=FALSE}
# Loading in metadata
Metadata <- read.csv("HEB114_Metadata.csv") 
    Metadata$Mouse <- as.character(Metadata$Mouse) # lil data cleaning

# Loading in endpoint data
Endpoint <- read.csv("HEB114_Endpoint.csv")
    Endpoint$Mouse <- as.character(Endpoint$Mouse) # lil data cleaning

# Making new col for cecal contents
Endpoint <- Endpoint %>% 
  mutate(Cecal_contents_g = Cecum_full_g - Cecum_empty_g)

# Joining metadata and endpoint data 
Endpoint <- full_join(Endpoint, Metadata, by = "Mouse")

# summary table
GI_summary <- Endpoint %>% 
  group_by(ABX, Hormone) %>%
  summarize(Average_SI_length = round(mean(SI_length), 2),
            Average_colon_length = round(mean(Colon_length), 2),
            Average_cecal_contents = round(mean(Cecal_contents_g, na.rm = TRUE), 2)
            ) # Note: na.rm = TRUE because we have one row with NA value that we want to ignore

# making abx and non abx sub data frames
Endpoint_nonabx <- Endpoint %>% 
  filter(ABX == "Non_ABX")

Endpoint_abx <- Endpoint %>% 
  filter(ABX == "ABX")
```

## 1.a. Overview
Our endpoint phenotypic data consists of the following measurements:

1. Small intestine length (cm)
2. Colon length (cm)
3. Cecal contents (g)

Here's an interactive table to take a look at the raw endpoint data: 
```{r Endpoint data table}
datatable(Endpoint)
```

Here's a summary table of the three endpoint phenotypic parameters across the six groups:
```{r Summary GI table}
datatable(GI_summary)
```

**For phenotypic data, our analyses will be broken down as follows:** 

1. Does Measurement A significantly differ between Hormone X and Saline within the same ABX group?
    + *e.g. does SI length significantly differ between ABX mice receiving saline and PGH?*
  
2. Does Measurement A significantly differ between mice with and without a typical gut microbiome receiving the same hormone?  
    + *e.g. does SI length significantly differ between ABX and Non-ABX mice receiving PL?*

## 1.b. Regressions of SI vs Colon

Before examining each individual metric, I was curious what the relationship is between colon and SI lengths. Do mice with larger SIs also tend to have larger colons? 
```{r Regressions of SI vs Colon}

# First I plotted this for all mice, coloring the dots by abx group and distinguishing hormone groups with shapes. 
colon_vs_si <- ggplot(data = Endpoint %>% drop_na(Colon_length, SI_length), aes(x = Colon_length, y = SI_length)) +
   geom_point(aes(shape = Hormone, color = ABX)) + 
    geom_smooth(method = "lm") + 
   labs(title = "Linear Regression of \n SI and Colon Length",
        x = "Colon length (cm)", y = "SI length (cm)") 

print(colon_vs_si)

#Is this a significant correlation?
colon_vs_si_lm <- lm(SI_length ~ Colon_length, data = Endpoint %>% drop_na(Colon_length, SI_length)) # running the lm

colon_vs_si_lm_summary <- summary(colon_vs_si_lm) # checking it out

# Making a function that prints a summary of my results of interest
  analyze_lm <- function(lm_summary) {
  # Extract p-value and Adjusted R-squared
  p_value <- lm_summary$coefficients[2, 4]  # p-value for the predictor
  adj_r_squared <- lm_summary$adj.r.squared # Adjusted R-squared
  
  # Determine significance
  significance <- ifelse(p_value < 0.05, "significant.", "not significant.")
  
  # Print the result
 return(paste("The linear model is", significance, 
               "p =", signif(p_value, 3), 
               "and R^2 =", signif(adj_r_squared, 3), "."))
  }
  
analyze_lm(colon_vs_si_lm_summary)

# Does this relationship differ based on ABX status?
colon_vs_si2 <- ggplot(data = Endpoint %>% drop_na(Colon_length, SI_length), aes(x = Colon_length, y = SI_length)) +
   geom_point(aes(shape = Hormone, color = ABX)) + 
    geom_smooth(method = "lm", aes(color = ABX)) + 
   labs(title = "Linear Regression of \n SI and Colon Length",
        x = "Colon length (cm)", y = "SI length (cm)") 
print(colon_vs_si2)
# it appears that it might hold true for abx mice but not for non abx mice...

# let's check statistically
# lm for abx mice
colon_vs_si_abx_lm <- lm(SI_length ~ Colon_length, data = Endpoint_abx %>% drop_na(Colon_length, SI_length)) # running model
colon_vs_si_abx_lm_summary <- summary(colon_vs_si_abx_lm) # summarize

cat("For the ABX mice, ", analyze_lm(colon_vs_si_abx_lm_summary)) # print results

# lm for non-abx mice
colon_vs_si_nonabx_lm <- lm(SI_length ~ Colon_length, data = Endpoint_nonabx %>% drop_na(Colon_length, SI_length)) # running model
colon_vs_si_nonabx_lm_summary <- summary(colon_vs_si_nonabx_lm) # summarize
cat(" For the Non-ABX mice, ", analyze_lm(colon_vs_si_nonabx_lm_summary)) # print results

# it doesn't reach significance in either group, but the pattern visually appears to be stronger in the ABX mice. 
```

Those results were not particularly interesting. 

## 1.c. Small Intestine Lengths Across Groups 

First let's just take a look at the data across all groups:
```{r Small Intestine Lengths Across Groups 1}
ggplot(data = Endpoint, aes(x = Hormone, y = SI_length, fill = Hormone)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = Hormone, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("ABX") + 
  scale_fill_manual(values = c("plum3", "aquamarine3", "goldenrod")) +
  labs(title = "SI Lengths by Hormone Exposure \n and ABX Status \n (Outliers Included)",
       x = "Hormone", y = "SI length (cm)") +
  scale_y_continuous(limits = c(30, 45)) +
  stat_compare_means(
    comparisons = list(
      c("PGH", "Saline"),  # Compare PGH to Saline
      c("PL", "Saline")    # Compare PL to Saline
    ),
    method = "t.test",     # Use t-test
    label = "p.format"
  )
```

NOTES:  

  + There are some outliers in this data. I'm going to filter to remove those points and replot. 
  + I think ideally I'd want to include the outliers in the visualization and just remove them from the stats, but I couldn't figure out how to do that. 
  + I'm running these stats as t-tests for now. Could change to wilcox later if I wanted / needed to.
  + The y-axes ranges look a bit wonky because I want them to be consistent across plots while leaving room for the labels. 

```{r Small Intestine Lengths Across Groups 2}

# Now I'm creating a function to remove outliers within each group
remove_outliers <- function(data, column, group) {
  data_clean <- data %>%
    group_by(!!sym(group)) %>%  # Group by the specified column
    mutate(
      Q1 = quantile(.data[[column]], 0.25),
      Q3 = quantile(.data[[column]], 0.75),
      IQR = Q3 - Q1,
      lower_bound = Q1 - 1.5 * IQR,
      upper_bound = Q3 + 1.5 * IQR,
      is_outlier = .data[[column]] < lower_bound | .data[[column]] > upper_bound
    ) %>%
    ungroup()
  
  outliers <- data_clean %>% filter(is_outlier)
  
  # Print a message
  if (nrow(outliers) > 0) {
    cat("Outliers detected and removed:", nrow(outliers), "outlier(s).\n")
  } else {
    cat("No outliers detected.\n")
  }
  
  # Return data without outliers
  return(data_clean %>% filter(!is_outlier) %>% select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound, -is_outlier))
}

# Apply the function to remove outliers from the Endpoint dataset
Endpoint_clean <- remove_outliers(Endpoint, "SI_length", "Group_Total")

ggplot(data = Endpoint_clean, aes(x = Hormone, y = SI_length, fill = Hormone)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = Hormone, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("ABX") + 
  scale_fill_manual(values = c("plum3", "aquamarine3", "goldenrod")) +
  labs(title = "SI Lengths by Hormone Exposure \n and ABX Status \n (Outliers Excluded)",
       x = "Hormone", y = "SI length (cm)") +
  scale_y_continuous(limits = c(30, 45)) +
  stat_compare_means(
    comparisons = list(
      c("PGH", "Saline"),  # Compare PGH to Saline
      c("PL", "Saline")    # Compare PL to Saline
    ),
    method = "t.test",     # Use t-test
    label = "p.format"
  )

```

*Takeaways:* *With outliers excluded,* there is a signif difference in SI length between PL and saline-exposed mice with a conventional gut microbiome, but not between those who received antibiotics. No other within-abx-group comparisons reached significance. When the data is left as-is, there are no significant differences between hormone groups. 

Now let's rearrange the data to be faceted by hormone group, comparing abx and non abx mice. 
```{r Small Intestine Lengths Across Groups 3}
# complete dataset
ggplot(data = Endpoint, aes(x = ABX, y = SI_length, fill = ABX)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = ABX, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("Hormone") + 
  scale_fill_manual(values = c("chartreuse4", "goldenrod")) +
  labs(title = "SI Lengths by Hormone Exposure \n and ABX Status \n (Outliers Included)",
       x = "Hormone", y = "SI length (cm)") +
  scale_y_continuous(limits = c(30, 45)) +
  stat_compare_means(
    comparisons = list(
      c("ABX", "Non_ABX")),
    method = "t.test",
    label = "p.format"
  ) 

# outlier removed dataset 
ggplot(data = Endpoint_clean, aes(x = ABX, y = SI_length, fill = ABX)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = ABX, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("Hormone") + 
  scale_fill_manual(values = c("chartreuse4", "goldenrod")) +
  labs(title = "SI Lengths by Hormone Exposure \n and ABX Status \n (Outliers Excluded)",
       x = "Hormone", y = "SI length (cm)") +
  scale_y_continuous(limits = c(30, 45)) +
  stat_compare_means(
    comparisons = list(
      c("ABX", "Non_ABX")),
    method = "t.test",
    label = "p.format"
  ) 
```
*Takeaways:* As expected, mice exposed to antibiotics have longer small intestines than those with a complete conventional gut microbiome. However, this difference only reaches significance in Saline and PGH mice. In PL mice, we approach but don't quite hit the 0.05 threshold. This is true regardless of the inclusion or exclusion of outliers.  

*** 

## 1.d. Colon Lengths Across Groups 
```{r Colon Lengths Across Groups 1}
ggplot(data = Endpoint, aes(x = Hormone, y = Colon_length, fill = Hormone)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = Hormone, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("ABX") + 
  scale_fill_manual(values = c("plum3", "aquamarine3", "goldenrod")) +
  labs(title = "Colon Lengths by Hormone Exposure \n and ABX Status \n (Outliers Included)",
       x = "Hormone", y = "Colon length (cm)") +
    scale_y_continuous(limits = c(5, 9)) +
  stat_compare_means(
    comparisons = list(
      c("PGH", "Saline"), 
      c("PL", "Saline") 
    ),
    method = "t.test",
    label = "p.format"
  )

```
*Takeaways:* No observed difference in colon length across hormone groups. I didn't bother to remove outliers from this data because it seemed clear to me that it wouldn't make any difference. 

Now let's rearrange to compare each hormone group by abx status:
```{r Colon Lengths Across Groups 2}
ggplot(data = Endpoint, aes(x = ABX, y = Colon_length, fill = ABX)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = ABX, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("Hormone") + 
  scale_fill_manual(values = c("chartreuse4", "goldenrod")) +
  labs(title = "Colon Lengths by Hormone Exposure \n and ABX Status \n (Outliers Included)",
       x = "Hormone", y = "Colon length (cm)") +
  scale_y_continuous(limits = c(5, 9)) +
  stat_compare_means(
    comparisons = list(
      c("ABX", "Non_ABX")),
    method = "t.test",
    label = "p.format"
  ) 
```

```{r Colon Lengths Across Groups 3, include=FALSE}
# outliers excluded
Endpoint_clean_2 <- remove_outliers(Endpoint, "Colon_length", "Group_Total")

ggplot(data = Endpoint_clean_2, aes(x = ABX, y = Colon_length, fill = ABX)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = ABX, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("Hormone") + 
  scale_fill_manual(values = c("chartreuse4", "goldenrod")) +
  labs(title = "Colon Lengths by Hormone Exposure \n and ABX Status \n (Outliers Excluded)",
       x = "Hormone", y = "Colon length (cm)") +
  scale_y_continuous(limits = c(5, 9)) +
  stat_compare_means(
    comparisons = list(
      c("ABX", "Non_ABX")),
    method = "t.test",
    label = "p.format"
  ) 

```
*Takeaways:* Colon lengths were generally longer in mice exposed to antibiotics in all hormone exposure groups, but this only reached significance in the PL-exposed mice. Note: I reran this excluding the one saline non-abx outlier, but it made no difference. The range of colon lengths within each group was also quite large, especially for the abx-exposed mice.  

*** 

## 1.e. Cecal Contents Across Groups 
```{r Cecal Contents Across Groups 1}

ggplot(data = Endpoint, aes(x = Hormone, y = Cecal_contents_g, fill = Hormone)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = Hormone, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("ABX") + 
  scale_fill_manual(values = c("plum3", "aquamarine3", "goldenrod")) +
  labs(title = "Cecal Mass by Hormone Exposure \n and ABX Status \n (Outliers Included)",
       x = "Hormone", y = "Cecal Mass (g)") +
  scale_y_continuous(limits = c(0.2, 1.8)) +
  stat_compare_means(
    comparisons = list(
      c("PGH", "Saline"), 
      c("PL", "Saline") 
    ),
    method = "t.test",
    label = "p.format"
  )

```
*Takeaways:* No observed difference in cecal mass across hormone groups. I didn't bother to remove outliers from this data because it seemed clear to me that it wouldn't make any difference. 

```{r Cecal Contents Across Groups 2}
ggplot(data = Endpoint, aes(x = ABX, y = Cecal_contents_g, fill = ABX)) + 
  geom_boxplot() +
  geom_jitter(aes(fill = ABX, alpha = 0.5), width = 0.2, height = 0) + 
  theme_bw() + 
  facet_wrap("Hormone") + 
  scale_fill_manual(values = c("chartreuse4", "goldenrod")) +
  labs(title = "Cecal Mass by Hormone Exposure \n and ABX Status \n (Outliers Included)",
       x = "Hormone", y = "Cecal Mass (g)") +
  scale_y_continuous(limits = c(0.2, 1.8)) +
  stat_compare_means(
    comparisons = list(
      c("ABX", "Non_ABX")),
    method = "t.test",
    label = "p.format"
  ) 
```
*Takeaways:* As expected, regardless of hormone exposure, mice exposed to antibiotics have much larger cecums than those with a conventional gut microbiome. 

